---
title: "Intro to R"
author: "Kyla McConnell"
output: html_document
---

# Welcome to the tidyverse

Welcome to the tidyverse! Tidyverse is a package -- really, a whole collection of packages -- that include a LOT of useful functions for all sorts of data analyzing, cleaning and "wrangling". 

To start, make sure you have tidyverse installed. You can do this either through the panel at the lower left hand corner (Packages tab) or by typing into the Console `install.packages("tidyverse")`. Then, call the tidyverse with a `library()` call and let's get started!

```{r}
library(tidyverse)
```

Tidyverse loads multiple key packages:
- dplyr -> for all sorts of data transformation and wrangling
- ggplot2 -> the best plotting package in R (and ever??)
- readr & tibble -> for reading in files to tibble format (improvements over R base data.frames)
- stringr -> for text transformations (removing trailing whitespaces, etc.)
- magrittr -> for pipes, more on this below

It also loads purr (for vectorized programming), forcats (improvements to factors), readxl (for reading Excel documents), lubridate (for working with dates/times) and more.

### The pipe %>% 

One of the most noticeable features of the tidyverse is the pipe %>% (keyboard shortcut: Ctr/Cmd + Shift + M) 

The pipe takes the item before it and feeds it to the following command as the first argument. Since all tidyverse (and some non-tidyverse) functions take the dataframe as the first function, this can be used to string together multiple functions. 

Load in the following dataset, which contains IKEA furniture items in Saudi Arabia and their prices (in Saudi Riyals)
```{r}
ikea <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-11-03/ikea.csv')
```

First, take a look at the dataset. You can do this with `head(ikea)` or try out using the pipe. 

```{r}
ikea %>% 
  head()
```

You see that this produces the exact same output as `head(ikea)`. Why would this be useful?

Compare the following lines of pseudocode, which would produce the same output:

A.
bop(
  scoop(
    hop(foo_foo, through = forest),
    up = field_mice
  ), 
  on = head
)


B.
foo_foo %>%
  hop(through = forest) %>%
  scoop(up = field_mice) %>%
  bop(on = head)
  
You can see that the version with the pipe is easier to read when more than one function is called on the same dataframe! 

### Selecting columns

A useful step in using pipes and tidyverse calls is the ability to *select* specific columns. That is, instead of writing `ikea$category` we can write:

```{r}
ikea %>% 
  select(category) %>% 
  head()
```

Here, I've piped the output to `head()` to reduce the amount of rows shown in the preview (don't do this if you want to work with the whole column!)

We can then use this column for further calculations, like counting the number of unique categories. Note here that instead of `length(unique(ikea$category))`, we use `count()` -- a tidyverse function that works better here
```{r}
ikea %>% 
  select(category) %>% 
  unique() %>% 
  count()
```





