---
title: "Exercises"
author: "Julia Müller"
date: "1 2 2021"
output: html_document
---

At the moment this is just a dump of exercises I took out of the main doc!

# Functions and arguments
Call the `say()` function including the arguments what, by, and what_color. Try at least a few different options for each of these and find your favorite look. Try using "random" with the argument "by" or "catfact" or "fortune" with the argument "what".
```{r}

```


# Reading in
Read in the files you'll need for the exercises (in the exercises folder). Have a look at the data you read in. Is the separator recognised correctly? Are all variables the correct type? If not, convert them.

# Piping 

1. We tried piping catdog_spr to `head()` above. Try it now with `tail()`. What does this show you? 
```{r}

```

2. Pipe the 'catdog_spr' dataset to `colnames()` What does this command do? (Bonus, pipe the columnn names on to the `nchar()` function and figure out what that does.)
```{r}

```


# Fixing code errors

All of the following code snippets produce errors. What have I done wrong? Correct each snippet.

```{r}
party_invites <- c(Tracey, Karen, Sandra)
```

```{r}
1, 4, 5 + 1
```

```{r}
max(price) #from the ikea tibble
```

```{r}
ikea %>% 
  filter(item = "BILLY")
```

```{r}
ikea %>% 
  select(starts_with(item))
```

```{r}
ikea %>% 
  mutate(price * 2 = new_price)
```

```{r}
sumary(ikea)
```

```{r}
42 + "5"
```


# filter

2. Reading times lower than 150ms are pretty unusual and are sometimes removed from self-paced reading data. You decide that you want to remove these response times from your data. Filter out all RTs below 150ms from the spr dataset and save your result.
```{r}
nrow(spr)

spr <- spr %>% 
  filter(RT >= 150)

nrow(spr)
```

3. You want to take a look at the results from Par_C only: Filter these data points but don't save the result. 
```{r}
spr %>% 
  filter(participant == "Par_C")
```

4. You're not interested in the practice items in the spr dataframe. Filter the dataset to include everything BUT the practice items and save your result.
```{r}
spr <- spr %>% 
  filter(cond != "practice")
```


# If-else

1. Create a new column that shows whether or not the word is in the first half of a sentence. If the word number is less than or equal to 3, this new column should read 'first half', if not, it should read 'second half'.
```{r}
spr %>% 
  mutate(sentence_half = ifelse(word_num <= 3, "first half", "second half"))
```

2. Create a new column that has the part of speech for a word. If the word is in following list, it should be labeled "content", otherwise, it should be labeled "function": c("practice", "sentence", "relative", "clause.", "cats.")
```{r}
spr %>% 
  mutate(POS = ifelse(word %in% c("practice", "sentence", "relative", "clause.", "cats."), "content", "function"))
```

3. Create a new column that shows how fast a word was read. First, find the mean reading time over the entire dataset. You can do this in the console most easily using the base-R syntax `mean(spr$RT)`. If the RT is greater than or equal to this number, your new column should read "slow", otherwise, it should read "fast". 
```{r}
mean(spr$RT)

spr %>% 
  mutate(fastness = ifelse(RT >= mean(spr$RT), "slow", "fast"))
```

# Group by, summarise

1. Starting with the spr dataset, group by participant and use summarize to find the max RT for each participant. 
```{r}
spr %>% 
  group_by(participant) %>% 
  summarize(max(RT)) %>% 
  ungroup()
```

2. With the same dataset, group by condition and return the mean RT for each condition. This is already a first step towards statistical analysis of your data! Do conditions look like they have similar or very different means?
```{r}
spr %>% 
  group_by(cond) %>% 
  summarize(mean(RT)) %>% 
  ungroup()
```


3. Summarize based on condition and count how many full sentences are in each condition. (Here you will realize that this example dataset doesn't make a ton of sense :D)
```{r}
spr %>% 
  group_by(cond) %>% 
  count(full_sentence) %>% 
  ungroup()

spr %>% 
  count(full_sentence)
```

